\LARGE \bf Face Recognition: Too Bias, or Not Too Bias?

### 2020-02-15 manuscript.tex
name changes to more generic filenames
wip
Updates from Overleaf

\usepackage{url}            % simple URL typesetting
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
%\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{tikz}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
\usepackage{arydshln}
\usepackage{todonotes}
\usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}

% \usepackage[small]{caption}
% \usepackage{subcaption}
% \usepackage{subfigure} 
\usepackage{sidecap}
\usepackage{wrapfig}
\usepackage[subrefformat=parens,labelformat=parens]{subcaption} %% provides subfigure
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{float}% If comment this, figure moves
\usepackage{geometry}

@@

\usepackage{array}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

@@

\newacronym{sdm}{SDM}{signal detection model}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{nmse}{NMSE}{Normalized Mean Square Error}
\newacronym{det}{DET}{Detection Error Trade-off}
\newacronym{tp}{TP}{true-positive}
\newacronym{fp}{FP}{false-positive}

@@

\newacronym{fpr}{FPR}{false-positive rate}
\newacronym{tpr}{TPR}{true-positive rate}

@@

\title{\LARGE \bf\vspace{-10mm}
Face Recognition: Too Bias, or Not Too Bias?
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. Classic signal detection theory revealed trends in the underlying score distribution across subgroups. Specifically, we show variations in the optimal scoring threshold varies for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. To download the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}
\
\input{sections/introduction.tex}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
\begin{center}
    \caption{\small{\textbf{Quantitative of human assessment.} Different human subgroups listed per row. Each column is the subgroup labeled. Note that each people are best within their subgroup, and second-best within the same subgroup but different gender. CF shows the least variation, but with the lowest accuracy. CM shows the best accuracy, but second to \gls{wm} in deviation from the mean. Thus, scores for males vary more than females.}}
    \label{tab:humsn-eval-results} 
     \vspace{-2mm}

\footnotesize
\scalebox{0.9}{
\begin{tabular}{l c c c c c}
        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
        CF &  \textbf{0.529}&  0.480&0.438&0.447 &0.474$\pm$0.041 \\
        CM & 0.456 & \textbf{0.504}  & 0.444 &0.362 &0.441$\pm$0.059 \\
        WF & 0.447 &0.438 &  \textbf{0.573}& 0.480 & 0.485$\pm$0.062 \\
        WM & 0.301&0.474 &  0.453 & \textbf{0.561} & 0.447$\pm$0.108 \\\midrule
        Avg &  0.433& 0.474&0.477 &0.463 &0.462$\pm$0.020\\
 \end{tabular}}
 \end{center}
 \vspace{-6mm}
\end{table}

\glsresetall
\section{CONCLUSION}
We introduce a new data set \gls{bfw} with eight subgroups balanced across gender and ethnicity. With this, and upon highlighting the challenges and shortcomings of grouping subjects as a single subset, we provide evidence that forming subgroups is meaningful, as the \gls{fr} algorithm rarely makes mistakes across subgroups. We trained Arc-Face net on MSCeleb, expecting that the results would suffer from bias because of the imbalanced train-set. Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the \gls{fpr} up to a factor of two, which is seemingly the cause of the frenzy about bias in \gls{fr} in main-stream media. Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out \gls{fpr}, and achieving a higher \gls{tpr}. We hypothesized that most humans grown amongst more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best. The focused research findings presented here, along with the public database included, are extendable in vast ways. Thus, we see this as the slither to a much larger problem of bias in ML.

s\bibliographystyle{ieee}
\bibliography{bias}

%\putbib
%\end{bibunit}
%\begin{bibunit}
%\input{supplemental.tex}
%\end{bibunit}

>>>>

\usepackage{url}            % simple URL typesetting
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
%\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{tikz}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
\usepackage{arydshln}
\usepackage{todonotes}
\usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{balance}
% \usepackage[small]{caption}
% \usepackage{subcaption}
% \usepackage{subfigure} 
\usepackage{sidecap}
\usepackage{wrapfig}
\usepackage[subrefformat=parens,labelformat=parens]{subcaption} %% provides subfigure
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{float}% If comment this, figure moves
\usepackage{geometry}

@@

% \usepackage{minipage}
\usepackage{array}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

@@

\newacronym{sdm}{SDM}{signal detection model}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{nmse}{NMSE}{Normalized Mean Square Error}
\newacronym{det}{DET}{Detection error trade-off}
\newacronym{tp}{TP}{true-positive}
\newacronym{fp}{FP}{false-positive}

@@

\newacronym{fpr}{FPR}{false-positive rate}

\newacronym{tpr}{TPR}{true-positive rate}

@@

\title{\LARGE \bf
Face Recognition: Too Bias, or Not Too Bias?
}

% \title{\LARGE \bf\vspace{-10mm}
% Face Recognition: Too Bias, or Not Too Bias?
% }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. We show variations in the optimal scoring threshold for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}
\
\input{sections/introduction.tex}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
\begin{center}
    \caption{\small{\textbf{Quantitative of human assessment.} Different human subgroups listed per row. Each column is the subgroup labeled. Note that each people are best within their subgroup, and second-best within the same subgroup but different gender. CF shows the least variation, but with the lowest accuracy. CM shows the best accuracy, but second to \gls{wm} in deviation from the mean. Thus, scores for males vary more than females.}}
    \label{tab:humsn-eval-results} 
    %  \vspace{-2mm}
\footnotesize
\scalebox{0.9}{
\begin{tabular}{l c c c c c}
        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
        CF &  \textbf{0.529}&  0.480&0.438&0.447 &0.474$\pm$0.041 \\
        CM & 0.456 & \textbf{0.504}  & 0.444 &0.362 &0.441$\pm$0.059 \\
        WF & 0.447 &0.438 &  \textbf{0.573}& 0.480 & 0.485$\pm$0.062 \\
        WM & 0.301&0.474 &  0.453 & \textbf{0.561} & 0.447$\pm$0.108 \\\midrule
        Avg &  0.433& 0.474&0.477 &0.463 &0.462$\pm$0.020\\
 \end{tabular}}
 \end{center}
%  \vspace{-7mm}
\end{table}

\glsresetall
\section{Conclusion}
We introduce a new data set \gls{bfw} with eight subgroups balanced across gender and ethnicity. With this, and upon highlighting the challenges and shortcomings of grouping subjects as a single subset, we provide evidence that forming subgroups is meaningful, as the FR algorithm rarely makes mistakes across subgroups. We used an \textit{off-the-shelf} Sphereface, hypothesizing this SOTA CNN suffers from bias because of the imbalanced train-set. Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the FPR up to a factor of two, which is seemingly the cause of the frenzy about bias in FR in main-stream media. Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out FPR, and achieving a higher TPR. We hypothesized that most humans grown amongst more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best. The focused research findings presented here, along with the public database included, are extendable in vast ways. Thus, we see this as the slither to a much larger problem of bias in ML.

% \vspace{-3mm}
\bibliographystyle{ieee}
\balance
\bibliography{references}

% \putbib
% \end{bibunit}
\begin{bibunit}
\input{supplemental.tex}
\end{bibunit}

### 2020-02-15 sections/analysis.tex
Updates from Overleaf

\begin{figure}[t!]
	\centering    
	\includegraphics[width=\linewidth]{figures/confusion.pdf}
		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others. Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic). This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
		\label{fig:confusion} 
		\vspace{-5mm}
\end{figure}

\section{Results and Analysis}
An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https:\/\/github.com\/clcarwin\/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped. The two features were fused by concatenation.

@@

% \vspace{-2mm}
% \vspace{-2pt}
\subsection{Score Analysis}
Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot. Notice that score distributions for imposters tends to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots. On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape). Asian Fem
Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups. A vast majority of errors occurs in intra-subgroup. It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate. The categories are, therefore, meaningful in the context of \gls{fr}.

@@

\begin{table}[t!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
 \vspace{-2mm}
\begin{center}
\begin{tabular}{l c c c c c}
     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
    \midrule
    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
\end{tabular}
\end{center}
\glsreset{tar}
\glsreset{far}
 \vspace{-3mm}

@@

Even for lower \gls{far}, there are notable improvements, often of the order of 1\%, which can be challenging to achieve when \gls{far} is near $\geq$90\%. More importantly, each subgroup has the desired \gls{fpr}, so that substantial differences in \gls{fpr} will remain unfounded. We further experimented using ethnicity estimators and using the ethnicity of both query and test image, which yielded similar results to those reported here (results not shown).

\begin{figure}[t!] 
	\centering    
	\includegraphics[width=\linewidth] {figures/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}. Accuracy scores shown as bar plots. Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
		\vspace{-5mm}
\end{figure} 
\vspace{-2mm}
% \vspace{-5pt}
\subsection{Human evaluation}
 Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully. One might expect that the most exposure to others would be within the same subgroup, and, therefore, would be best at labeling their own. Secondarily, they would be best at labeling images of the same ethnicity, but opposite gender. Our findings concur. Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex. Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-to-best having the same ethnicity and opposite gender. On average, subgroups are comparable at labeling images.

>>>>

\begin{figure}[t!]
	\centering    
	\includegraphics[width=\linewidth]{figures/confusion.pdf}
		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others. Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic). This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
		\label{fig:confusion} 
\end{figure}

\section{Results and Analysis}
An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped. The two features were fused by average pooling (\ie 512 D).

@@

\subsection{Score analysis}
Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot. Notice that score distributions for imposters tends to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots. On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape). Asian Fem
Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups. A vast majority of errors occurs in intra-subgroup. It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate. The categories are, therefore, meaningful in the context of \gls{fr}.

@@

\begin{table}[t!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
%  \vspace{-2mm}
\begin{center}
\begin{tabular}{l c c c c c}
     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
    \midrule
    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
\end{tabular}
\end{center}
\glsreset{tar}
\glsreset{far}
%  \vspace{-3mm}

@@

Even for lower \gls{far}, there are notable improvements, often of the order of 1\%, which can be challenging to achieve when \gls{far} is near $\geq$90\%. More importantly, each subgroup has the desired \gls{fpr}, so that substantial differences in \gls{fpr} will remain unfounded. We further experimented using ethnicity estimators and using the ethnicity of both query and test image, which yielded similar results to those reported here.

\begin{figure}[t!] 
	\centering    
	\includegraphics[width=\linewidth] {figures/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}. Accuracy scores shown as bar plots. Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
% 		\vspace{-1mm}
\end{figure} 
% \vspace{-4mm}
% \vspace{-5pt}
\subsection{Human evaluation}
 Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully. Subjects of a subgroup likely have mostly been exposured to others of the same. Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but other gender. Our findings concur. Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex. Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-to-best having the same ethnicity and opposite gender. On average, subgroups are comparable at labeling images.

### 2020-02-15 sections/introduction.tex
Updates from Overleaf

\begin{figure}[t!]
\centering
         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{figures/montage.pdf}
\caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectfully. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
\label{fig:avg-faces}
% \vspace{-3pt}
\end{figure}

@@

    \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
    \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
    Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
    \gls{dp}~\cite{demogPairs}     & CASIA-W, VGG\&VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
    \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
    \end{tabular}
    \label{tab:compared}
    \vspace{-5mm}
\end{table*}

@@

        Acc$@t_o$ & 0.942 &0.952 &0.968 &0.961 &0.960 &0.962 & 0.974 &0.982 & 0.963 $\pm$ 0.012\\
        \bottomrule
    \end{tabular}}
    \glsunset{af}
    }
    \glsunset{af}
    \glsunset{am}
    \glsunset{bf}
    \glsunset{bm}
    \glsunset{if}
    \glsunset{im}
    \glsunset{wf}
    \glsunset{wm}
    \vspace{-12pt}
\end{table*}

The contributions in this paper are as follows:
\begin{enumerate}
    \item We built a balanced dataset as a proxy to measure verification performance per subgroup.
    \item We analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
    \item We showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    \item We conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
\end{enumerate}

>>>>

\begin{figure}[t!]
\centering
         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=\linewidth]{figures/montage.pdf}
\caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectfully. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
\label{fig:avg-faces}
\end{figure}

@@

    \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
    \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
    Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
    \gls{dp}~\cite{demogPairs}     & CASIA-W, VGG\&VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
    \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
    \end{tabular}
    \label{tab:compared}
    % \vspace{-4mm}
\end{table*}

@@

        Acc$@t_o$ & 0.942 &0.952 &0.968 &0.961 &0.960 &0.962 & 0.974 &0.982 & 0.963 $\pm$ 0.012\\
        \bottomrule
    \end{tabular}}
    \glsunset{af}
    }
    \glsunset{af}
    \glsunset{am}
    \glsunset{bf}
    \glsunset{bm}
    \glsunset{if}
    \glsunset{im}
    \glsunset{wf}
    \glsunset{wm}
    % \vspace{-12pt}
\end{table*}

The contributions in this paper are as follows:
\begin{enumerate}
    \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
    \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
    \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
\end{enumerate}

### 2020-02-15 sections/relatedworks.tex
Updates from Overleaf

Most similar to our work are the recent efforts in~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face}. A common factor of these efforts are claims of insufficient data supply to support studies on bias in \gls{fr}, and the introduction of new metadata on demographics for image sets parsed out of existing collections. Specifically, \cite{demogPairs} curated a set of faces based on racial demographics (\ie Asian, Black, and White) called \gls{dp}, while~\cite{srinivas2019face} honed in on adults versus children called gls{itwcc}. Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest. Besides, the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender. Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.

>>>>

Most similar to our work are the recent efforts in~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face}. A common factor of these efforts are claims of insufficient data supply to support studies on bias in \gls{fr}, and the introduction of new metadata on demographics for image sets parsed out of existing collections. Specifically, \cite{demogPairs} curated a set of faces based on racial demographics (\ie Asian, Black, and White) called \gls{dp}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}. Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest. Besides, the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender. Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.
